\documentclass[preprint,12pt]{elsarticle}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{float}

\journal{Infrared Physics \& Technology}

\newcommand{\cmark}{\checkmark}
\newcommand{\xmark}{--}

\begin{document}

\begin{frontmatter}

\title{UAV-TAlign-12K: Benchmarking Scene-Level Reliability in\\ Cross-FOV UAV Infrared-Visible Alignment}

\author[anon]{Anonymous Authors}
\affiliation[anon]{organization={Anonymous submission}}

\begin{abstract}
Reliable UAV infrared-visible alignment is important for aerial inspection, nighttime monitoring,
and multimodal thermal sensing, yet existing resources provide limited support for evaluating
real captured cross-FOV alignment at the scene level. We introduce UAV-TAlign-12K, which contains
6,039 real-world UAV infrared-visible pairs across 15 scenes and provides an integrity-checked
6,037-pair official evaluation split. The dataset spans day, night, and low-light
capture, grayscale and pseudocolor thermal rendering, and wide, zoom, and standard UAV views. We
further define a manifest-fixed, label-free benchmark that separates pairwise homography evidence
from scene-level retained alignment and reports condition-aware and risk--coverage diagnostics.
Evaluation with seven off-the-shelf baselines, spanning classical, infrared-visible-specific, and
modern learned methods, shows that pairwise evidence is broadly available: RIFT2, raw MINIMA, and
RoMa produce homographies for all 6,037 evaluated pairs, while LoFTR and XoFTR remain
near-saturated. As the benchmark reference framework, UAV-TAlign aggregates frame evidence
through reliability-guided progressive selection, robust consensus, and QA-aware verification. It produces
scene homographies for all 15 scenes and retains 9 under the strict canonical gate. Five-fold
leave-frame-out analysis yields a 1.39-pixel median scene-consensus drift, while controlled
translation, rotation, and scale perturbations produce monotonic degradation of the joint QA signal.
Together, the
dataset, protocol, and reference framework establish a reproducible benchmark for studying when
cross-FOV UAV infrared-visible alignment is reliable enough to retain.
\end{abstract}

\begin{highlights}
\item UAV-TAlign-12K provides 6,039 real cross-FOV UAV infrared-visible pairs.
\item Seven baselines establish broad pairwise homography availability.
\item UAV-TAlign produces 15/15 scene transforms and retains 9 at the strict gate.
\item Resampling and controlled perturbations validate the reliability protocol.
\end{highlights}

\begin{keyword}
infrared-visible dataset \sep UAV imagery \sep multimodal registration \sep alignment benchmark \sep scene-level reliability
\end{keyword}

\end{frontmatter}

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_dataset_protocol}
\input{sections/04_method}
\input{sections/05_experimental_setup}
\input{sections/06_results}
\input{sections/07_discussion}
\input{sections/08_conclusion}

\bibliographystyle{elsarticle-num}
\bibliography{references}

\end{document}
