% !TEX root = ../paper.tex

\section{Related Work}

\subsection{Classical Image Registration}

Classical image registration pipelines typically rely on hand-crafted local features followed by
robust geometric fitting. Representative methods built on SIFT, SURF, ORB, BRISK, KAZE, or AKAZE
remain attractive because they are simple, interpretable, and do not require task-specific
training
\cite{lowe2004sift,bay2008surf,rublee2011orb,leutenegger2011brisk,alcantarilla2012kaze,alcantarilla2013akaze}.
In cross-modal settings, the appearance gap between visible and thermal imagery can reduce
descriptor repeatability, especially when thermal frames contain low-texture or homogeneous
regions. These methods nevertheless provide a strong geometric reference point for understanding
which parts of the registration problem are addressed by local feature extraction and which parts
require higher-level scene reasoning.

These methods remain important baselines in our study because they represent the classical
registration view of the problem: stable local features followed by robust homography fitting.
Robust model estimation remains central in these pipelines, from RANSAC and multiple-view geometry
to more recent variants such as MAGSAC and MAGSAC++
\cite{fischler1981ransac,hartley2004multiple,barath2019magsac,barath2020magsacpp}. Our results
show that real UAV infrared-visible scenes benefit from moving beyond a single-pair formulation toward
scene-level evidence aggregation and reliability control.

\subsection{Deep Pairwise Matching for Multimodal Registration}

Recent dense and semi-dense matching methods have significantly improved pairwise geometric
correspondence estimation. Learned local features have evolved from trainable detectors and
descriptors such as SuperPoint, LF-Net, D2-Net, R2D2, DISK, S2DNet, HardNet, L2-Net, SOSNet,
GeoDesc, ContextDesc, ASLFeat, ALIKE, and Key.Net
\cite{detone2018superpoint,ono2018lfnet,dusmanu2019d2net,revaud2019r2d2,tyszkiewicz2020disk,germain2020s2dnet,mishchuk2017hardnet,tian2017l2net,tian2019sosnet,luo2018geodesc,luo2019contextdesc,luo2020aslfeat,zhao2023alike,barroso2019keynet},
while match filtering and end-to-end correspondence reasoning have progressed through
SuperGlue, NCNet, Patch2Pix, COTR, LoFTR, ASpanFormer, LightGlue, DKM, RoMa, Efficient LoFTR,
OmniGlue, MINIMA, and XoFTR
\cite{sarlin2020superglue,rocco2018ncnet,zhou2021patch2pix,jiang2021cotr,sun2021loftr,chen2022aspanformer,lindenberger2023lightglue,edstedt2023dkm,edstedt2024roma,wang2024efficientloftr,jiang2024omniglue,ren2025minima,tuzcuoglu2024xoftr}.
In particular, these methods are better at recovering useful correspondences in low-texture
regions, repetitive structures, or substantial appearance shifts. Broad reviews of this transition
from classical pipelines to trainable matching, together with practical benchmarking studies,
provide additional context
\cite{ma2021survey,liao2024survey}.

Our work builds directly on this progress by treating modern matchers as high-capacity evidence
generators. The empirical picture in UAV-TAlign-12K is that strong off-the-shelf multimodal
matchers and infrared-visible classical baselines already provide very high pairwise availability.
This moves the central question from
correspondence generation alone to the conversion of frame-level estimates into stable and reliable
scene-level alignment under realistic UAV operating conditions.

\subsection{Visible-Thermal Registration and Evaluation}

Visible-thermal registration has been studied under a range of assumptions, from same-view or
near-same-view alignment to broader multimodal correspondence learning. Radiation-variation
insensitive matching methods such as RIFT and deep alignment or stitching formulations such as
CLKN and UDIS indicate the breadth of existing registration formulations
\cite{li2020rift,chang2017clkn,nie2021udis}. Many existing settings emphasize relatively
controlled viewpoint changes or pairwise registration outcomes. By contrast, practical UAV
infrared-visible data frequently combine cross-FOV capture, cross-resolution sensing, thermal rendering
variation, and low-light conditions.
These factors make scene-level reliability a more central concern than in many standard
visible-thermal evaluation setups.

The most closely related UAV alignment benchmark direction is represented by synthetic
transform-supervised visible-infrared alignment datasets and trainable baselines, such as the
recent TGRS UAVMatch benchmark and transformer-based alignment network
\cite{gao2025uavmatch}. That line provides controlled affine or homography supervision, train/test
splits at large scale, and geometric metrics such as RMSE, AEE, and PCK. UAV-TAlign-12K is
designed for the complementary real-data setting: it fixes an integrity-checked captured
cross-FOV evaluation split, evaluates off-the-shelf infrared-visible evidence generators, and
studies how frame-level homographies become retained scene-level alignment products under a
label-free reliability protocol.

UAV-TAlign-12K extends this evaluation landscape by packaging real cross-FOV infrared-visible scenes
together with scalable label-free diagnostics. The protocol measures scene consistency and relative
alignment quality, allowing the study to focus on reliability-controlled scene-level alignment
under practical UAV acquisition conditions rather than isolated pairwise correspondence counts.

\subsection{UAV Multimodal Datasets and the Remaining Gap}

UAV multimodal datasets have supported progress in detection, tracking, segmentation, and broader
cross-modal perception, spanning aligned pedestrian and automotive benchmarks as well as
visible-thermal detection and tracking benchmarks such as KAIST, FLIR, LLVIP, M3FD, DroneVehicle,
VTUAV, LasHeR, and Anti-UAV
\cite{hwang2015kaist,flir2018thermal,jia2021llvip,liu2022tardal,sun2022dronevehicle,zhang2022vtuav,li2022lasher,jiang2023antiuav}.
These resources have primarily advanced downstream perception tasks, while scene-level
infrared-visible alignment places additional emphasis on reliability-aware geometric decision making.
In particular, a practical registration pipeline must decide not only how to align frames, but also
when the scene-level evidence is strong enough to trust the resulting transform.

This is the gap our paper targets. Complementary to new matcher architectures, our primary focus is
a real captured dataset and benchmark protocol for cross-FOV UAV infrared-visible alignment. The
collection and label-free protocol make pairwise evidence, retained scene transforms, and
condition-dependent reliability directly studyable. UAV-TAlign is provided as a secondary
scene-level reference method for instantiating this benchmark rather than as a replacement for
matcher-centric research.
