% !TEX root = ../paper.tex

\section{Experimental Setup}

\subsection{Baselines}

We evaluate both classical and learning-based baselines under an explicit off-the-shelf setting.
All learning-based methods use public pretrained weights or public pretrained model interfaces and
are evaluated without training or adaptation on UAV-TAlign-12K. This choice keeps the comparison
focused on practical deployability rather than dataset-specific adaptation.
Unified inference wrappers standardize model loading, image I/O, homography fitting, and output
schemas while preserving the public pretrained baseline parameters and settings.

The baseline suite covers classical keypoint pipelines, an infrared-visible-specific method, and
modern learned matchers. The classical pairwise references are SIFT and AKAZE, each followed by
robust homography fitting. RIFT2 is evaluated through its Python implementation under a
resize-aware setting. The modern pairwise baselines are Kornia LoFTR, RoMa outdoor full,
XoFTR-640, and raw MINIMA. LoFTR
uses Kornia's public outdoor-pretrained model, while raw MINIMA uses the MINIMA RoMa branch as a
direct pairwise baseline. Reliability-guided progressive evidence selection, robust
scene consensus, and QA-aware verification are activated only in the full UAV-TAlign framework
\cite{sun2021loftr,edstedt2024roma,tuzcuoglu2024xoftr,ren2025minima}.

\subsection{Shared Geometric Fitting}

For the pairwise methods, the common downstream step is homography fitting from the predicted
correspondences. After each matcher returns point pairs and optional confidence values, the
shared evaluator estimates a homography with a robust backend and then records a unified outcome
category. This unified post-processing is important because it makes the pairwise baseline
table comparable at the level of usable registration outcomes rather than raw matcher outputs
alone.

The shared evaluator uses a robust homography backend across the integrated pairwise baselines:
\begin{itemize}
\item robust fitting method: USAC-MAGSAC;
\item reprojection threshold: 3.0;
\item confidence: 0.999;
\item maximum iterations: 10000;
\item spatial coverage grid: $4 \times 4$.
\end{itemize}

RIFT2 is independently integrated and uses its frozen evaluated setting:
\texttt{max\_dim=1200}, \texttt{npt=2000}, \texttt{patch\_size=64}, Lowe ratio 0.95, and
USAC-MAGSAC with a 5.0-pixel reprojection threshold. Its fitting statistics are therefore reported
as method-specific complementary diagnostics rather than treated as a scalar cross-method rank.

\subsection{Main-Table Baseline Choices}

The main-table baseline choices are as follows. Raw MINIMA uses the MINIMA RoMa backend
with the \texttt{minima\_roma} checkpoint family and the large/full RoMa branch. RoMa uses the
public outdoor full model. XoFTR uses the 640-resolution visible-thermal checkpoint family as the
default main-table setting. Kornia LoFTR (pretrained="outdoor") uses the public pretrained outdoor
model from Kornia.

\subsection{Evaluation Details}

The LoFTR baseline is evaluated under a resize-aware inference setting. The longest
image dimension used for matching is limited to 1200, and automatic mixed precision is enabled.
This profile preserves the public pretrained model while supporting consistent full-subset
evaluation.

We additionally conduct two controlled protocol validations on the frozen formal outputs. First,
five-fold leave-frame-out analysis partitions the accepted frame hypotheses deterministically,
recomputes scene consensus after withholding one fold, and measures the resulting grid displacement
from the full-evidence transform. Second, a synthetic sensitivity test applies translation,
rotation, and scale perturbations at nominal severities of 2, 5, 10, 20, and 40 pixels to three
representative frames per scene. It records the response of edge overlap and gradient-NCC without
changing the canonical decision rule.

Journal-scale experiments were conducted in isolated Python environments on an NVIDIA RTX 4090
workstation. The main environment used Python 3.10 with PyTorch 2.0.1, torchvision 0.15.2, CUDA
11.8, OpenCV 4.11.0, and Kornia 0.6.11. All runs used the fixed official manifest, fixed public
weights or model interfaces, isolated output roots, and read-only raw dataset directories.
