% !TEX root = ../paper.tex

\section{Discussion}

\subsection{What UAV-TAlign-12K Reveals}

The main contribution of UAV-TAlign-12K is a captured dataset and protocol that expose a gap hidden
by pairwise-only reporting. Modern matchers provide abundant frame-level evidence on the official
split, yet the strict scene-level criterion retains only a subset of scenes. The benchmark
therefore separates two practically different capabilities: producing a pairwise homography and
deciding whether multi-frame evidence supports a trustworthy scene transform.

The protocol evidence supports this dataset-centered reliability framing. Retained and QA-filtered scenes show
structured differences in accepted ratio, severe outlier behavior, and baseline-relative QA
signals, indicating that the canonical criterion captures scene-level reliability beyond pairwise
matcher availability alone. This gives the evaluation a clear diagnostic role: it tests when
abundant pairwise evidence becomes a trustworthy scene-level transform. The canonical scene
criterion is therefore designed for conservative retention of scene transforms rather than for
maximizing unconditional scene acceptance.

The controlled validations reinforce this interpretation from two independent directions.
Leave-frame-out recomputation keeps the scene consensus close to its full-evidence estimate, while
known translation, rotation, and scale perturbations progressively reduce the joint QA signal.
Together, these results connect multi-frame consensus stability with sensitivity to controlled
geometric degradation.

UAV-TAlign serves as a reference implementation for this second evaluation unit. Its evidence
selection, robust consensus, and QA-aware verification demonstrate one reproducible way to convert
pairwise hypotheses into a retained scene product; the benchmark remains compatible with future
matchers and alternative scene-level aggregation strategies.

The multi-seed random-selection supplement further strengthens the design rationale for
deterministic evidence selection. On the accepted 8-scene subset, random selection spans 5/8 to
7/8 retained scenes across seeds, but the variation is concentrated in a few borderline scenes
rather than in large shifts of accepted-frame support. Deterministic-even selection therefore
provides a reproducible default operating policy, while robust scene consensus and QA-aware
verification define the retained-product decision layer.

\subsection{Operating Scope and Extensibility}

UAV-TAlign-12K targets captured UAV scenes where a dominant planar or far-field transform is a
useful operating model. Its 15 scenes span complementary illumination, thermal rendering, view, and
scene-family conditions, establishing a focused evaluation resource for cross-FOV homography
alignment. The same manifest and evaluation interfaces can be extended to additional flight
altitudes, sensor configurations, and geometric models.

The label-free QA proxies provide scalable scene-consistency diagnostics, while the fixed canonical
gate and the non-trained reliability ranking expose complementary retained-product and operating-
profile views. This modular design supports direct combination with task-specific annotations,
alternative scene geometries, learned quality estimators, and downstream thermal-sensing tasks.
