% !TEX root = ../paper.tex

\section{Conclusion}

We introduced UAV-TAlign-12K, a benchmark-scale UAV infrared-visible alignment resource with
6,039 synchronized pairs, an integrity-checked 6,037-pair evaluation split, and a multi-level
protocol spanning pairwise evidence, scene consensus, and reliability--coverage analysis. Seven
representative baselines show that infrared-specific and modern matchers already provide highly
available pairwise homographies across captured cross-FOV conditions.

UAV-TAlign converts this evidence into 15 scene transforms. It combines progressive scheduling,
geometry-calibrated frame weighting, robust consensus, and cross-modal verification.
At the canonical operating point, it identifies nine high-confidence products covering 74.4\% of
the evaluated pairs. Leave-frame-out resampling and controlled perturbations further support the
stability and geometric sensitivity of the scene diagnostics. These results establish scene-scale
evidence consolidation and confidence-aware operating analysis as central evaluation dimensions
for cross-FOV UAV infrared-visible alignment.

\section*{Data Availability}

UAV-TAlign-12K, the official evaluation manifest, metadata, integrity report, and evaluation
scripts have been prepared for public release. The official evaluation split contains 6,037
integrity-checked RGB-infrared pairs from a 6,039-pair UAV infrared-visible alignment collection.
The dataset, owned by Guangxi University, will be released under the Creative Commons Attribution
4.0 International license, with code and evaluation materials provided through a public
repository. Review access can be provided during peer review. The official manifest hash is
provided for reproducible evaluation.
